---
title: "covid_daily"
author: "Max Davis"
date: "3/25/2020"
output: html_document
---

install.packages("readr")
install.packages("tidyr")
install.packages("ggplot2")
install.packages("data.table")
install.packages("dplyr")
install.packages("lubridate")
install.packages("curl")
install.packages("knitr")




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(tidyr)
library(ggplot2)
library(data.table)
library(dplyr)
library(lubridate)
library(curl)
library(RColorBrewer)

```

Load daily data from European Center for Disease Control:

```{r}
covid_cumul <- (fread("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv"))
```
Change to data.frame and change date format to system date format.
```{r}

covid_cumul <- data.frame(covid_cumul)
```

```{r}
covid_cumul$dateRep <- format(as.Date(covid_cumul$dateRep, format="%d/%m/%Y"),"%Y-%m-%d")
```

Add feature:Deaths over Cases. deathRatio 

```{r}
  covid_cumul$deathRatio <-
    covid_cumul$deaths / covid_cumul$cases
```

Create new feature: Cases as percentage of population. case_percentage.

```{r}
covid_cumul$case_percentage <-
    covid_cumul$cases / covid_cumul$popData2019
```

Store system date as today

```{r}
today <- Sys.Date()
```
Show only data from today.

covid_new <-
    covid_cumul %>%
    filter(deaths >= 1) %>%
    filter(dateRep == today)


View(covid_new)



Number of new deaths


sum(covid_new$deaths)

Filter out wanted countries from cumulative data

```{r}
covid_hotspots <- 
  covid_cumul %>%
  filter(geoId == "EC" | geoId == "CO" | geoId == "HU" | geoId == "JO" | geoId == "US" | geoId == "CZ" | geoId == "UK" | geoId == "CN" | geoId == "NO" | geoId == "CA") 
```         

Filter out only new data from the hotspot data
    
```{r}
 covid_hotspots_new <-
  covid_hotspots %>%
    filter(deaths >= 1) %>%
    filter(dateRep == today)
```

PLot new deaths reported in hotspots:
```{r}
 ggplot(covid_hotspots, aes(x = geoId, y = deaths))+
  geom_point() +
  ggtitle("Deaths Reported Today in Hotspot Countries")
  
```

Data Frame with only US data from mid march:

```{r}
covid_US <- 
  covid_cumul %>%
    filter(geoId == "US") %>%
    filter(dateRep >= "2020-03-14")

```

Data Frame with hotspot data from mid march

```{r}
covid_march_compare <-
    covid_hotspots %>%
    filter(dateRep >= "2020-03-15")
```
Data frame with hotspot data from after July 4
```{r}
covid_july_compare <- 
    covid_hotspots %>%
    filter(dateRep >= "2020-07-04")
```


Plot change in US cases over time, compare with other hotspots

```{r}
covid_march_compare %>%
  ggplot(aes(x = as.Date(dateRep), y = cases)) +
  geom_line(aes(color = countriesAndTerritories)) +
  ggtitle("New Cases over Time in Top Countries") +
  labs(x = "Date", y = "Reported Cases", color = "Country")
```

Visualize ratio of deaths to cases reported.
Plot deathRatio to compare countries.

```{r}
covid_march_compare %>%
  ggplot(aes(x = as.Date(dateRep), y = deathRatio)) +
  geom_line(aes(color = countriesAndTerritories)) +
  ggtitle("Ratio of deaths to new cases") +
  labs(x = "Date", y = "Ratio of Deaths to new Cases", color = "Country")
```





Plot the percentage of cases to population since May 15. Still plots march? 


```{r}
covid_july_compare %>%
  ggplot(aes(x = as.Date(dateRep), y = case_percentage)) +
  geom_line(aes(color = countriesAndTerritories)) +
  ggtitle("Cases as Percentage of Population Over Time") +
  labs(x = "Date", y = "Cases as Percentage of Population", color = "Country")
```


Looking at China






















